---
title: "team_challenge"
output: pdf_document
---

```{r}
library(e1071)
library()
```
# Data Cleaning

We began our analysis by first loading the training data set and then examine the predicators.

```{r}
data <- read.csv('data/cs-training.csv')
colnames(data)
```

## Portion of defaulted
```{r}
mean(data$SeriousDlqin2yrs)
```


We then check each feature with summary() and see which of these features have null data and how many.

```{r}
data_col <- colnames(data)
for(i in 2:12){
  print(data_col[i])
  print(summary(data[,i]))
}
```
This suggests that only monthly income and number of dependents have missing data, which we would later either fill in or drop.


```{r}
train_complete<- complete.cases(data)
data_drop<- cbind(data,train_complete)
data_drop<- subset(data_drop,data_drop$train_complete == TRUE)
```
^These code above give us the dataframe with incompleted observations dropped 

```{r}
data_median<-read.csv("/Users/sophiasun/Desktop/154/GiveMeSomeCredit/cs-training.csv",header= TRUE,sep=",")

for (i in 1:150000){
   if (is.na(data_median$MonthlyIncome[i])== TRUE) {
     data_median$MonthlyIncome[i] = 5400
   }
}

for (i in 1:150000){
   if (is.na(data_median$NumberOfDependents[i])== TRUE) {
     data_median$NumberOfDependents[i] = 0
   }
  }

```
^These code gave us data_median which uses the respective medians to fill in NAs in MonthlyIncome and NumberOfDependents
